---
title: "League of Legends 2022 Match Data Analysis"
output: 
  html_notebook:
    toc: true
    toc_float: true
    code_folding: hide
    df_print: paged
---

# Introduction

## Preamble

The plan here is to use this dataset, kindly provided by Tim Sevenhuysen over at Oracle's Elixir, to practice what I'm learning on the CodeClan Data Analysis bootcamp. Along with the git repository, it'll allow me to work on this through the weeks of the course and have a detailed history of adjustments via git, plus hopefully have a pretty interesting workbook to show off once I'm done.

### Things I've learned

- Basics of `dplyr`
- Creating functions
- Some RMarkdown
- Looping functions to do some automated table creation

## Data

### What is this data?

The data I'm reading in is an import of all League of Legends e-sports matches across pretty close to, if not all, the leagues operating in the world in 2022. Until the end of the year, this data will be updated regularly, so I want to make my work agnostic to the amount of data presented -- the way the data is laid out in the imported .csv is extremely unlikely to change, but more rows will be added over time. This allows me to just redownload a newer version of the data whenever I feel like it and everything should still work.

### How is the data broken down in the import?

Every match is broken down into 12 rows --

-- 1 row for each of the 5 players on each team
<br>
-- 1 row for for each team 

Some of the statistics are assigned to the teams, but not individual players.

For example, which team secured an objective first, or total counts on things like towers that haven't been tracked at a player level. In other cases it serves as an aggregate of the group of 5 players, i.e. total kills, total deaths, etc.

It's worth bearing this in mind for analysis, as if you try to examine the objectives secured by an individual player, you won't have any luck!

Every match has been assigned a `data completeness`, which refers in most cases to the depth of reporting: incomplete columns typically don't have breakdowns by 10 or 15 minutes, analyse play in the jungle, track kill-streaks like pentakills, etc. The general data will be there, but the granularity will not. 

**Any analysis of these data-points will probably exclude incomplete matches.**

# First Steps

## Libraries

Currently using 

- `tidyverse`
- `janitor`
- `snakecase`

```{r libraries,warning=FALSE, message=false}
library(tidyverse)
library(janitor)
library(snakecase)
```


## Data Import

First step is to read in the .csv

```{r import_data}
global_data_master <- read_csv("data/2022_match_data.csv")
```

## Clean-up

A whole bunch of these columns can be booleans. `result` is a binary check for win/loss, `playoffs` is a binary check for end of split playoffs, all the first blood columns are to do with just identifying involvement, as is `firstdragon`, `firstherald`, `firstbaron`, and the three metrics surrounding first tower. 

Going to rename `result` as `winner` for better clarity and snake_case all the column headers. Hopefully at some point I figure out a cleaner way to do this.

```{r format_global_data}
global_data <- global_data_master %>% 
    mutate(across(c(result,
                    playoffs,
                    firstblood, 
                    firstbloodkill, 
                    firstbloodassist,
                    firstbloodvictim,
                    firstdragon,
                    firstherald,
                    firstbaron,
                    firsttower,
                    firstmidtower,
                    firsttothreetowers), ~as.logical(.x))) %>% 
    rename(c("winner" = "result", 
             "game_id" = "gameid",
             "data_completeness" = "datacompleteness",
             "participant_id" = "participantid",
             "player_name" = "playername",
             "player_id" = "playerid",
             "team_name" = "teamname",
             "team_id" = "teamid",
             "ban_1" = "ban1",
             "ban_2" = "ban2",
             "ban_3" = "ban3",
             "ban_4" = "ban4",
             "ban_5" = "ban5",
             "game_length" = "gamelength",
             "team_kills" = "teamkills",
             "team_deaths" = "teamdeaths",
             "double_kills" = "doublekills",
             "triple_kills" = "triplekills",
             "quadra_kills" = "quadrakills",
             "penta_kills" = "pentakills",
             "team_kpm" = "team kpm",
             "c_kpm" = "ckpm",
             "elemental_drakes" = "elementaldrakes",
             "dragons_unknown" = "dragons (type unknown)",
             "damage_to_champions" = "damagetochampions",
             "damage_share" = "damageshare",
             "damage_taken_per_minute" = "damagetakenperminute",
             "damage_mitigated_per_minute" = "damagemitigatedperminute",
             "wards_placed" = "wardsplaced",
             "wards_killed" = "wardskilled",
             "control_wards_bought" = "controlwardsbought",
             "vision_score" = "visionscore",
             "total_gold" = "totalgold",
             "earned_gold" = "earnedgold",
             "earned_gpm" = "earned gpm",
             "earned_gold_share" = "earnedgoldshare",
             "gold_spent" = "goldspent",
             "total_cs" = "total cs",
             "minion_kills" = "minionkills",
             "monster_kills" = "monsterkills",
             "monster_kills_own_jungle" = "monsterkillsownjungle",
             "monster_kills_enemy_jungle" = "monsterkillsenemyjungle",
             "gold_at_10_mins" = "goldat10",
             "xp_at_10_mins" = "xpat10",
             "cs_at_10_mins" = "csat10",
             "opp_gold_at_10" = "opp_goldat10",
             "opp_xp_at_10" = "opp_xpat10",
             "opp_cs_at_10" = "opp_csat10",
             "gold_diff_at_10_mins" = "golddiffat10",
             "xp_diff_at_10_mins" = "xpdiffat10",
             "cs_diff_at_10_mins" = "csdiffat10",
             "kills_at_10_mins" = "killsat10",
             "assists_at_10_mins" = "assistsat10",
             "deaths_at_10_mins" ="deathsat10",
             "opp_kills_at_10" = "opp_killsat10",
             "opp_assists_at_10" = "opp_assistsat10",
             "opp_deaths_at_10" = "opp_deathsat10",
             "gold_at_15_mins" = "goldat15",
             "xp_at_15_mins" = "xpat15",
             "cs_at_15_mins" = "csat15",
             "opp_gold_at_15" = "opp_goldat15",
             "opp_xp_at_15" = "opp_xpat15",
             "opp_cs_at_15" = "opp_csat15",
             "gold_diff_at_15_mins" = "golddiffat15",
             "xp_diff_at_15_mins" = "xpdiffat15",
             "cs_diff_at_15_mins" = "csdiffat15",
             "kills_at_15_mins" = "killsat15",
             "assists_at_15_mins" = "assistsat15",
             "deaths_at_15_mins" ="deathsat15",
             "opp_kills_at_15" = "opp_killsat15",
             "opp_assists_at_15" = "opp_assistsat15",
             "opp_deaths_at_15" = "opp_deathsat15",
             "first_blood" = "firstblood", 
             "first_blood_kill" = "firstbloodkill",
             "first_blood_assist" = "firstbloodassist",
             "first_blood_victim" = "firstbloodvictim",
             "first_dragon" = "firstdragon",
             "first_herald" = "firstherald",
             "first_baron" = "firstbaron",
             "first_tower" = "firsttower",
             "first_mid_tower" = "firstmidtower",
             "first_to_three_towers" = "firsttothreetowers"
             ))
```

## Basic Breakdowns

Let's have a look at some basic information about the data.

```{r generate_global_stats}
games <- global_data %>% 
  filter(!is.na(game_id)) %>% 
  group_by(game_id) %>% 
  summarise() %>% 
  count() %>% 
  pull()

leagues <- global_data %>% 
  filter(!is.na(league)) %>% 
  group_by(league) %>% 
  summarise() %>% 
  count() %>% 
  pull()

players <- global_data %>% 
  filter(!is.na(player_name)) %>% 
  group_by(player_name) %>% 
  summarise() %>% 
  count() %>% 
  pull()

player_ids <- global_data %>% 
  filter(!is.na(player_id)) %>% 
  group_by(player_id) %>% 
  summarise() %>% 
  count() %>% 
  pull()

global_stats <- tibble(games, leagues, players, player_ids)

rm(player_ids,players,leagues,games)

global_stats
```

There's a mismatch between the number of players and their IDs. Are some names attributed to multiple IDs?

```{r find_duplicate_players}
global_data %>% 
  filter(!is.na(player_name), !is.na(player_id)) %>% 
  select(player_name, player_id) %>% 
  group_by(player_name) %>% 
  filter(n_distinct(player_id) > 1) %>% 
  distinct() %>% 
  summarise(id_count = n()) %>% 
  arrange(player_name)
```

```{r generate_global_tibbles}
global_wins <- global_data %>% 
  filter(participant_id == 100 & winner == TRUE | participant_id == 200 & winner == TRUE) %>% 
  group_by(team_name) %>% 
  summarise(wins = n()) %>% 
  arrange(desc(wins))

global_losses <- global_data %>% 
  filter(participant_id == 100 & winner == FALSE | participant_id == 200 & winner == FALSE) %>% 
  group_by(team_name) %>% 
  summarise(losses = n()) %>% 
  arrange(desc(losses))

global_table <- global_wins %>% 
  inner_join(global_losses, by = "team_name") %>% 
  mutate(played = rowSums(across(c("wins", "losses")))) %>% 
  filter(team_name != "unknown team") %>% 
  arrange(desc(wins))

global_table
```

```{r global_cleanup, include=FALSE}
rm(global_wins, global_losses)
```

## Functions

Next I'll make a series of functions to do some legwork for me.

NOTE: In a normal environment I'd define these way up top, but this file is documenting my learning process!

`get_league()` is really basic and let's me create a table containing just a specific league

`list_league()` looks for the unique league names and sticks them in a list

`get_league_table()` produces a Wins/Losses/Played league table for a specified league

`get_stats()` produces a table with some quick summary stats for a given league

`get_roster()` produces a table containing the roster for the specified team, including position and games played

`get_matches()` produces a vector containing the `game_id` of every game the specified team has played.

`get_game_history()` produces a game history for the specified team.

`get_match_history()` produces a match history for the specified team. NOTE -- "match history" is an aggregation of every individual "game" into set series so that, for example, the outcome of a Best of 5 is evaluated rather than the individual games in the series.

```{r functions}

get_league <- function(x_league, match_data){
  
  # Return tibble of specified league

    match_data %>%
      filter(league == str_to_upper((x_league)))
}

list_leagues <- function(match_data){
  
  # Return list of all leagues in match_data
  
  match_data %>% 
    group_by(league) %>% 
    summarise() %>% 
    distinct() %>% 
    pull()
}

get_league_table <- function(x_league){
  
  # Create string to reference a specific tibble
  league_var = to_snake_case(paste0(x_league,"_data"))
  league_wins <- list()
  league_losses <- list()
  
  # Use get() to call this object by name
  league_wins <- get(league_var) %>% 
    
    filter(participant_id == 100 & winner == TRUE | participant_id == 200 & winner == TRUE) %>% 
    group_by(team_name) %>% 
    summarise(wins = n()) %>% 
    arrange(desc(wins))
  
  league_losses <- get(league_var) %>% 
    filter(participant_id == 100 & winner == FALSE | participant_id == 200 & winner == FALSE) %>% 
    group_by(team_name) %>% 
    summarise(losses = n()) %>% 
    arrange(desc(losses))
  
  return(as.data.frame(league_wins) %>% 
    inner_join(as.data.frame(league_losses), by = "team_name") %>% 
    mutate(played = rowSums(across(c("wins", "losses")))))
}

get_stats <- function(x_league){
  
  # Create string to reference a specific tibble
  league_var = to_snake_case(paste0(x_league,"_data"))
  
  league_stats <- get(league_var)
  
  games <- league_stats %>% 
    filter(!is.na(game_id)) %>% 
    group_by(game_id) %>% 
    summarise() %>% 
    count() %>% 
    pull()
  
  teams <- league_stats %>% 
    filter(!is.na(team_name)) %>% 
    group_by(team_name) %>% 
    summarise() %>% 
    count() %>% 
    pull()
  
  players <- league_stats %>% 
    filter(!is.na(player_name)) %>% 
    group_by(player_name) %>% 
    summarise() %>% 
    count() %>% 
    pull()
  
  player_ids <- league_stats %>% 
    filter(!is.na(player_id)) %>% 
    group_by(player_id) %>% 
    summarise() %>% 
    count() %>% 
    pull()
  
  return(tibble(games, teams, players, player_ids))
}

get_roster <- function(x_team, match_data){
  
  match_data %>% 
  filter(!is.na(player_name), team_name == x_team) %>% 
  select(position, player_name, game_id) %>% 
  group_by(player_name) %>% 
  summarise(position, games = n_distinct(game_id)) %>% 
  distinct() %>% 
  ungroup() %>% 
  select(position, player_name, games) %>% 
  arrange(factor(position, levels = c("top", "mid", "bot", "sup", "jng")), player_name)
}

get_matches <- function(x_team, match_data){
  
  match_data %>% 
  select(game_id, team_name) %>% 
  group_by(game_id) %>%
  filter(team_name == x_team) %>% 
  select(game_id) %>% 
  distinct() %>% 
  pull()
}

get_game_history <- function(x_team, match_data){
  
  # I could use get_matches() here...
  
  team_games <- match_data %>% 
  select(game_id, team_name) %>% 
  group_by(game_id) %>%
  filter(team_name == x_team) %>% 
  select(game_id) %>% 
  distinct() %>% 
  pull()
  
  match_data %>% 
  select(game_id, date, league, split, playoffs, game, side, team_name, winner) %>% 
  filter(game_id %in% team_games, team_name != x_team) %>%
  distinct() %>% 
  mutate(winner = !winner) %>% 
  arrange(date, game)
}

get_match_history <- function(x_team, match_data){

  # I could use get_matches() here...
  team_games <- match_data %>% 
  select(game_id, team_name) %>% 
  group_by(game_id) %>%
  filter(team_name == x_team) %>% 
  select(game_id) %>% 
  distinct() %>% 
  pull()
  
  # I could use get_game_history() here...
  game_history <- match_data %>% 
  select(game_id, date, league, split, playoffs, game, team_name, winner) %>% 
  filter(game_id %in% team_games, team_name != x_team) %>%
  distinct() %>% 
  mutate(winner = !winner) %>% 
  arrange(date, game)
  
  game_history %>% 
  select(date, league, split, playoffs, game, team_name, winner) %>% 
  mutate(date = as.Date(date)) %>% 
  group_by(date) %>%
  count(team_name, league, playoffs, split, winner) %>% 
  pivot_wider(names_from = winner, values_from = n) %>% 
  rename(wins = `TRUE`, losses = `FALSE`) %>% 
  mutate(wins = coalesce(wins, 0), losses = coalesce(losses, 0),
         winner = if_else(wins > losses, TRUE, FALSE))
}
```

```{r AUTOMATION fetch all leagues, eval=FALSE, include=FALSE}

# Produce tibbles for all leagues in the format "[league]_data"

leagues_vector <- list_leagues(global_data)

for(i in 1:length(leagues_vector)){
  
   assign(to_snake_case(paste0(leagues_vector[i],"_data")), 
          get_league(leagues_vector[i], global_data))
}

rm(leagues_vector)
```

```{r AUTOMATION create all league tables, eval=FALSE, include=FALSE}
# Create win/loss/played tibbles for every league in the format "[league]_table"

leagues_vector <- list_leagues(global_data)

for(i in 1:length(leagues_vector)){
  
   assign(to_snake_case(paste0(leagues_vector[i],"_table")), 
          get_league_table(leagues_vector[i]))
}

rm(leagues_vector)
```

# Analysis of European (LEC) data

## Basics {.tabset}

Let's have some fun with the European data. The league we want to look at is the LEC.

```{r create Europe-only (LEC) data}
lec_data <- get_league("lec", global_data)
lec_table <- get_league_table("lec")
```

### League Table

```{r LEC League Table}
lec_table
```

### Summary

```{r LEC stats}
get_stats("lec")
```


## Team Compositions {.tabset}

Roster Sizes

```{r team sizes, message=FALSE}
lec_data %>% 
  select(team_name, player_name) %>% 
  filter(!is.na(team_name), !is.na(player_name)) %>% 
  group_by(team_name) %>% 
  summarise(no_players = n_distinct(player_name))
```

### Astralis

```{r Astralis roster, warning=FALSE, message=FALSE}
get_roster("Astralis", lec_data)
```

### Excel Esports

```{r Excel Esports roster, warning=FALSE, message=FALSE}
get_roster("Excel Esports", lec_data)
```

### Fnatic

```{r Fnatic roster, warning=FALSE, message=FALSE}
get_roster("Fnatic", lec_data)
```

### G2 Esports

```{r G2 Esports roster, warning=FALSE, message=FALSE}
get_roster("G2 Esports", lec_data)
```

### MAD Lions

```{r MAD Lions roster, warning=FALSE, message=FALSE}
get_roster("MAD Lions", lec_data)
```

### Misfits Gaming

```{r Misfits Gaming roster, warning=FALSE, message=FALSE}
get_roster("Misfits Gaming", lec_data)
```

### Rogue

```{r Rogue roster, warning=FALSE, message=FALSE}
get_roster("Rogue", lec_data)
```

### SK Gaming

```{r SK Gaming roster, warning=FALSE, message=FALSE}
get_roster("SK Gaming", lec_data)
```

### Team BDS

```{r Team BDS roster, warning=FALSE, message=FALSE}
get_roster("Team BDS", lec_data)
```

### Team Vitality

```{r Team Vitality roster, warning=FALSE, message=FALSE}
get_roster("Team Vitality", lec_data)
```

## Match History {.tabset}

As we move forward, I'm going to deepen the analysis on just "Rogue" and "G2 Esports", but thanks to all the functions I'm developing it would be very easy to repeat over other teams if I wished.

As a point of interest -- a regular league year is split into two seasonal "Splits" -- Spring and Summer, with a playoff tournament at the end of each to decide overall winners.

As you can see, these two teams played each other in the playoffs in both splits, with G2 winning Spring and Rogue winning Summer

### Rogue {.tabset}

#### Game History {.tabset}

##### Both splits

```{r Rogue game history}
get_game_history("Rogue", lec_data) %>% 
  select(-league)
```

##### Spring split

```{r Rogue Spring game history}
get_game_history("Rogue", lec_data) %>% 
  filter(split == "Spring") %>% 
  select(-c(league, split))
```

##### Summer split

```{r Rogue Summer game history}
get_game_history("Rogue", lec_data) %>% 
  filter(split == "Summer") %>% 
  select(-c(league, split))
```

#### Match History {.tabset}

##### Both splits

```{r Rogue match history}
get_match_history("Rogue", lec_data) %>% 
  select(-league)
```

##### Spring split

```{r Rogue Spring match history}
get_match_history("Rogue", lec_data) %>% 
  filter(split == "Spring") %>% 
  select(-c(league, split))
```

##### Summer split

```{r Rogue Summer match history}
get_match_history("Rogue", lec_data) %>% 
  filter(split == "Summer") %>% 
  select(-c(league, split))
```

### G2 Esports {.tabset}

#### Game History {.tabset}

##### Both splits

```{r G2 game history}
get_game_history("G2 Esports", lec_data) %>% 
  select(-league)
```
##### Spring split

```{r G2 spring game history}
get_game_history("G2 Esports", lec_data) %>% 
  filter(split == "Spring") %>% 
  select(-c(league, split))
```

##### Summer split

```{r G2 summer game history}
get_game_history("G2 Esports", lec_data) %>% 
  filter(split == "Summer") %>% 
  select(-c(league, split))
```

#### Match History {.tabset}

##### Both splits

```{r G2 match history}
get_match_history("G2 Esports", lec_data) %>% 
  select(-league)
```
##### Spring split

```{r G2 Spring match history}
get_match_history("G2 Esports", lec_data) %>% 
  filter(split == "Spring") %>% 
  select(-c(league, split))
```

##### Summer split

```{r G2 Summer match history}
get_match_history("G2 Esports", lec_data) %>% 
  filter(split == "Summer") %>% 
  select(-c(league, split))
```
